\chapter{Introduction}

\section{Scope and Objectives}
This guidebook documents the step-by-step flow used in the ICIIS workshop: starting from RTL (Verilog), performing synthesis and netlist generation, Static Timing Analysis (STA), placement and power planning, and finishing with routing and Clock-Tree Synthesis (CTS).

The primary objective is to provide a comprehensive workflow for converting a Register Transfer Level design into a GDSII layout using specific demonstration designs.

\section{How to Use This Guide}
Each chapter contains an ``Installation'' subsection with required tools and environment setup, followed by step-by-step instructions and expected outputs.

\section{Installation Overview}
To follow this workshop, the following general prerequisites are required:
\begin{itemize}
    \item \textbf{General Tools:} TeX live (for documentation), git.
    \item \textbf{Open-Source EDA Tools:} Icarus Verilog, GTKWave, Yosys, OpenSTA, OpenROAD, OpenLane.
    \item \textbf{Commercial Tools (Optional/Reference):} Synopsys Design Compiler, PrimeTime, and IC Compiler II (used for comparison or specific visualizations like the IO package view if open tools are insufficient).
\end{itemize}

Detailed installation instructions for specific tools are provided at the beginning of their respective chapters.

\section{Workshop File Layout}
The workshop materials are organized in the \texttt{RTL2GDS\_WS\_Files/} directory. The bundle contains the source RTL, synthesis scripts, and technology files referenced throughout the chapters:

\begin{lstlisting}[language=bash]
RTL2GDS_WS_Files/
|-- rtl/
|   |-- design6_comb.v          # Combinational design variant
|   |-- design6_seq.v           # Sequential design (main target)
|   |-- tb_design6.v            # Testbench
|   |-- design6_sim             # Compiled simulation (generated)
|   `-- design6.vcd             # Waveform output (generated)
|-- synth/
|   |-- run_synth.tcl           # Yosys synthesis script
|   `-- design6_netlist.v       # Gate-level netlist (generated)
|-- sta/
|   |-- design6.sdc             # Timing constraints
|   `-- run_sta.tcl             # OpenSTA script
|-- openroad/
|   |-- pdn_config.tcl          # Power distribution config
|   |-- run_placement.tcl       # Placement flow script
|   |-- run_routing.tcl         # Routing flow script
|   |-- design6_placed.def      # Placed design (generated)
|   |-- design6_final.def       # Routed design (generated)
|   |-- design6.drc             # DRC report (generated)
|   `-- design6.guide           # Routing guide (generated)
`-- lib/
    |-- sky130_fd_sc_hd__tt_025C_1v80.lib  # Liberty timing
    |-- sky130_fd_sc_hd__tt_025C_1v80.db   # Synopsys DB
    |-- sky130_fd_sc_hd.tlef               # Technology LEF
    |-- sky130_fd_sc_hd_merged.lef         # Standard cell LEF
    |-- sky130_fd_sc_hd.tf                 # Tech file
    `-- alib-52/                           # Compiled library
\end{lstlisting}

\noindent The technology libraries for the SkyWater SKY130 HD process are included in \texttt{RTL2GDS\_WS\_Files/lib/}:
\begin{itemize}
    \item \texttt{sky130\_fd\_sc\_hd\_\_tt\_025C\_1v80.lib} (Liberty timing/power)
    \item \texttt{sky130\_fd\_sc\_hd.tlef} and \texttt{sky130\_fd\_sc\_hd\_merged.lef} (technology + standard cell LEFs)
    \item \texttt{sky130\_fd\_sc\_hd.tf} plus auxiliary files such as \texttt{sky130\_fd\_sc\_hd\_\_tt\_025C\_1v80.db}
\end{itemize}